Penalized linear mixed models for structured genetic data

### 2021-03-15 main.tex
Sec 2.3, 3.8, 5 - in progress

\pb{What is the significance of this Mandt paper? Isn't there a huge literature on GLMMs? \citet{mandt2017sparse} generalize the LMM modeling paradigm to binary phenotypes in the interest of classification. They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference. }

@@

All analyses and simulations were performed using R version 3.6.3. The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}. As described previously, LMM-lasso-ggmix requires specialized software, but is implemented in the R package \texttt{ggmix} \citep{bhatnagar2020simultaneous}. Standard lasso software such as \texttt{glmnet} can be used to fit the LMM-lasso-Rakitsch method as well, although this requires a fair amount of preprocessing and setting up.  To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files. pb{Is this sentence important? Why do we care what parameterization Rakitsch used? Although \citet{rakitsch2013lasso} parameterize their model in terms of the variance components $\sigmass$ and $\delta := \sigmaee / \sigmass$, \texttt{penalizedLMM} implements the LMM-lasso-Rakitsch method using an equivalent but more interpretable parameterization in terms of $\eta$ and $\sigmaee$.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD. \pb{Mention imputation capabilities?}

@@

\pb{How do you feel about combining Sections 5.1 + 5.2, and Figures 5 + 6?  Each section seems really short to me.}

@@

\pb{I'm confused -- this formula does NOT require sign to be correct.  What's right, the formula or the text?}

Selection accuracy was evaluated using selection precision. All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} \ne 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly. Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_05.png}
\caption{\pb{I feel like we should stop this at 200.  The 100-400 range dominates this plot and what I get from this figure is: all the methods are equally bad.} Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc}
\end{figure}

@@

%------------------------------%
\subsection{PC-lasso and LMM-lasso increase genome-wide estimation accuracy}
%------------------------------%

@@

\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.}
The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso. However, PC-lasso performed somewhat better than LMM-lasso in this case study, in contrast to most of the simulation experiments. This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure. Thus, its structure is well-suited for PC-lasso. This finding suggests that while LMM-lasso may be more robust to complex structures, PC-lasso remains a strong competitor when the population structure is coarse.

>>>>

\pb{What is the significance of this Mandt paper? Isn't there a huge literature on GLMMs?}\anna{ \citet{mandt2017sparse} generalize the \textbf{high-dimensional, penalized} LMM modeling paradigm to binary phenotypes in the interest of classification. They do so by thresholding the outcome variable through a Probit likelihood with additional sparsity constraints and approximate Bayesian inference. }

@@

All analyses and simulations were performed using R version 3.6.3. The ordinary lasso and PC-lasso methods were implemented using \texttt{glmnet} \citep{glmnet}. As described previously, LMM-lasso-ggmix requires specialized software, but is implemented in the R package \texttt{ggmix} \citep{bhatnagar2020simultaneous}. Standard lasso software such as \texttt{glmnet} can be used to fit the LMM-lasso-Rakitsch method as well, although this requires a fair amount of preprocessing and setting up. To make this process easier, we developed an R package \texttt{penalizedLMM}, available at the github repository \url{https://github.com/areisett/penalizedLMM}, to allow seamless, user-friendly fitting of penalized LMMs based on genotype matrices or PLINK files. \anna{\texttt{penalizedLMM} includes preprocessing functions for \anna{SNP and sample-level filtering and} genotype imputation using \textbf{imputation rules - how to describe this} where possible, or else the Hardy-Weinberg Equilibrium value.} \texttt{penalizedLMM} additionally implements the elastic net, and the nonconvex penalties, MCP and SCAD. \anna{refine filtering, imputation functions}

@@

\pb{How do you feel about combining Sections 5.1 + 5.2, and Figures 5 + 6?  Each section seems really short to me.}
\anna{Fine with combining Sections 5.1 + 5.2 - I'm not sure what you mean by combine Figures 5 + 6. Keep one and not the other? Combine into a panel with parts (a) and (b)?}

@@

Selection accuracy was evaluated using selection precision. \anna{All $\beta > 0$ and we calculate precision = $\#\{ \beta > 0 \cap \widehat{\beta} > 0  \}/\#\{ \widehat{\beta} \ne 0 \}$ in order to ensure SNPs were not only selected, but their directions estimated correctly.} Figure~\ref{fig:gw_prc} displays each method's precision for the corresponding number of selected SNPs.
\begin{figure}[H]
\centering
\includegraphics[scale = 0.9]{figures/figure_05.png}
\caption{Precision for each method and number of selected SNPs from the Empirical genome-wide analysis. Single replication case study with $\eta = 0.5, \xi = 0.8$.}
\label{fig:gw_prc}
\end{figure}

@@

@@

\pb{I did change this paragraph a bit; I feel that we need to acknowledge and be honest about the fact that PC-lasso outperformed LMM-lasso here.} \anna{agreed}
The results of our genome-wide analysis case study broadly agree with the findings of our simulation experiments, in particular that PC-lasso and LMM-lasso are able to substantially reduce selection bias compared to the ordinary lasso. However, PC-lasso performed somewhat better than LMM-lasso in this case study, in contrast to most of the simulation experiments. This is perhaps to be expected as the sample consisted of three distinct racial groups with little admixture, and its RRM (Supplementary Figure~\ref{fig:empirical}) demonstrates a rather coarse population structure. Thus, its structure is well-suited for PC-lasso. This finding suggests that while LMM-lasso may be more robust to complex structures, PC-lasso remains a strong competitor when the population structure is coarse.
